Supplementary Material for ``Evaluating infectious disease forecasts with allocation scoring rules''

### 2023-12-22 alloscore_manuscript/supplement_short.tex
few more tweaks

A key feature of loss functions used in the decision theoretic definition of quantiles and related scoring rules such as the
CRPS and the WIS (see ...), as well as the allocation loss function presented in this work, is the presence of a \emph{shortage}:
the amount $\max\{0,y-x\}$ by which a resource demand
$y$ exceeds a supply decision variable $x$, which, for convenience, we write as $(y-x)_{+}$. In particular, a quantile at
probability level $\alpha$ of a distribution $F$ on $\mathbb{R}^1$ (which we assume to have a well-defined density $f(x)$)
is a Bayes act for the loss function
\[
l(x,y) = Cx + L(y-x)_{+}
\]
where $\alpha = 1-C/L$ and $C$ and $L$ can be interpreted as the cost per unit of a resource (such as medicine) and the loss
incurred when a unit of demand (such as illness) cannot be met due to the shortage $(y-x)_{+}$.  This follows because a
Bayes act, as a minimizer of $\Ex_F[l(x,Y)]$, must also be a vanishing point of the derivative
\begin{align}
\dby{}{x} \Ex_F\left[l(x,Y)\right] &= \Ex_F\left[\dby{}{x}l(x,Y)\right] \nonumber\\
&= C + L\Ex_F\left[\dby{}{x}(Y-x)_+\right] \nonumber\\
&= C - L\Ex_F\left[\mathbf{1}\{Y > x\}\right] \nonumber\\
&= C + L(F(x) - 1), \label{eqn:q_deriv}
\end{align}
so that $1-C/L = F(x)$.
The formula $\dby{}{x}\Ex_F\left[(Y-x)_+\right] = F(x) - 1$ for the derivative of the shortage,
used above in \eqref{eqn:q_deriv}, can be obtained from an application of the ``Leibniz Rule'':
\begin{align}
	\frac{d}{dx} \Ex_F [(Y-x)_{+}] &= \frac{d}{dx} \int_{x}^{\infty} (y-x) f_Y(y)dy \nonumber\\
	&= \int_{x}^{\infty} \frac{d}{dx}(y-x) f_Y(y)dy - (x-x) f_Y(x) = -\int_{x}^{\infty} f_Y(y)dy = F(x)-1. \label{eqn:shortage_deriv}
\end{align}
Note that more care is required when $F$ does not have a density.
We will also use this result below in deriving the Bayes act for the allocation loss.

@@

We find it helpful now to formally distinguish between two decision making problems. The first is the public health decision maker's allocation problem where the task is to select an allocation $x$, with the allocation loss $s_A(x, y) = \sum_{i=1}^N L \cdot \max(0, y_i - x_i)$ as described in section \ref{sec:methods.detailed.specific_allocation} of the main text.
The second is the forecaster's reporting problem where the task is to select parameter values $\theta$ to report. The forecaster's loss is given by $s_R(\theta, y) = s_A(x^{G_\theta}, y)$, where $x^{G_\theta}$ is the Bayes act for the allocation problem under the distribution $G_\theta$. In words, the loss associated with reporting $\theta$ is equal to the loss associated with taking the Bayes allocation corresponding to the distribution $G_\theta$.

@@

We define
$$G^{\star}(F) := \argmin_{G \in \mathcal{P}_{\mathrm{dfq}}} E_{F}[S_A(G, Y)],$$
the distribution corresponding to the optimal submission for the forecaster with predictive distribution $F$.
Because the distribution $G^q(F)$ typically will not yield allocations that are optimal under the distribution $F$, in general $G^q(F)$ and $G^\star(F)$ are different distributions.

>>>>

A key feature of loss functions for decision problems used to define quantiles and related scoring rules such as the
CRPS and the WIS (see e.g., \cite{gneiting2011quantiles}, \cite{jose2009evaluating}, 
and \cite{royset2022optimization}, sections 1.C and 3.C), as well as the allocation loss function presented in this work, 
is the presence of a \emph{shortage}:
the amount $\max\{0,y-x\}$ by which a resource demand
$y$ exceeds a supply decision variable $x$, which, for convenience, we write as $(y-x)_{+}$. In particular, a quantile at
probability level $\alpha$ of a distribution $F$ on $\mathbb{R}^1$ (which we assume to have a well-defined density $f(x)$)
is a Bayes act for the loss function
\[
l(x,y) = Cx + L(y-x)_{+}
\]
where $\alpha = 1-C/L$ and $C$ and $L$ can be interpreted as the cost per unit of a resource (such as medicine) and the loss
incurred when a unit of demand (such as illness) cannot be met due to the shortage $(y-x)_{+}$.  This follows because a
Bayes act, as a minimizer of $\Ex_F[l(x,Y)]$, must also be a vanishing point of the derivative
\begin{align}
\dby{}{x} \Ex_F\left[l(x,Y)\right] &= \Ex_F\left[\dby{}{x}l(x,Y)\right] \nonumber\\
&= C + L\Ex_F\left[\dby{}{x}(Y-x)_+\right] \nonumber\\
&= C - L\Ex_F\left[\mathbf{1}\{Y > x\}\right] \nonumber\\
&= C + L(F(x) - 1), \label{eqn:q_deriv}
\end{align}
so that $1-C/L = F(x)$.
The formula $\dby{}{x}\Ex_F\left[(Y-x)_+\right] = F(x) - 1$ for the derivative of the shortage,
used above in \eqref{eqn:q_deriv}, can be obtained from an application of the ``Leibniz Rule'':
\begin{align}
	\frac{d}{dx} \Ex_F [(Y-x)_{+}] &= \frac{d}{dx} \int_{x}^{\infty} (y-x) f_Y(y)dy \nonumber\\
	&= \int_{x}^{\infty} \frac{d}{dx}(y-x) f_Y(y)dy - (x-x) f_Y(x) = -\int_{x}^{\infty} f_Y(y)dy = F(x)-1. \label{eqn:shortage_deriv}
\end{align}
Note that more care is required when $F$ does not have a density.
We will also use this result below in deriving the Bayes act for the allocation loss.

@@

We find it helpful now to formally distinguish between two decision making problems. The first is the public health decision maker's allocation problem where the task is to select an allocation $x$, with the allocation loss $s_A(x, y) = \sum_{i=1}^N L \cdot \max(0, y_i - x_i)$ as described in section \ref{sec:methods.detailed.specific_allocation} of the main text.
The second is the forecaster's reporting problem where the task is to select parameter values $\theta$ to report. The forecaster's loss is given by 
\begin{align}
s_R(\theta, y) = s_A(x^{G_\theta}, y), \label{eqn:forecaster_theta_loss}
\end{align}
where $x^{G_\theta}$ is the Bayes act for the allocation problem under the distribution $G_\theta$. In words, the loss associated with reporting $\theta$ is equal to the loss associated with taking the Bayes allocation corresponding to the distribution $G_\theta$.

@@

We define
$$G^{\star}(F) := \argmin_{G \in \mathcal{P}_{\mathrm{dfq}}} E_{F}[S_A(G, Y)].$$
Since $S_R$ is defined as the Bayes scoring rule for the forecaster's loss \eqref{eqn:forecaster_theta_loss}, $G^{\star}(F)$ coincides
with $G_{\theta^{\star}(F)}$, the distribution in $\mathcal{P}_{\mathrm{dfq}}$ given by the 
optimal submission parameters $\theta^{\star}(F)$ for the forecaster with predictive distribution $F$.
In general, $G^q(F)$ and $G^\star(F)$ will be different distributions:  matching $F$ at specific quantiles
does not require $G^q(F)$ to match $F$ at the quantiles for $\tau^{F,K}$ (c.f. \eqref{eqn:quantiles-sum-to-K}), 
which would be necessary for it to share $x^F$ as an optimal allocation.
